---
title: "Tutorial 10: Linear Optimization"
author: "BT1101 Student. REPLACE WITH YOUR NAME"
date: 'Due by 15/04/2022 8:00 AM'
output: html_document
---

## Submission Instructions

- Select `output: html_document`.
- Include all code chunks, so include `echo=TRUE` in all chunks.
- Replace the placeholder text, "Type your answer here.", with your own.
- Submit *only* the required question for grading (Part 2: Submission). You can delete everything else for that submission. Remember to include any `library('package_name')` statements that you'll need to run your code and future reproduction. 
- Rename your R Markdown file `T[X]_[MatricNumber].rmd`, and the output will automatically be `T[X]_[MatricNumber].html`. 
- Submit your both R Markdown file (.rmd) and HTML (.html) to Luminus for tutorial assignments (upload to Luminus under the correct Submission Folder). We shall do the same for practical exam.
- **It is important to be able to code and produce your Rmarkdown output file *independently*.** You are responsible for de-bugging and programming in the practical exam.

## Preparation

```{r load-libraries, echo=TRUE, warning = FALSE, message = FALSE}
# load required packages
# install any package below if it's first time loaded in your computer.
library(lpSolve)
```

Please use the following table template for both Parts 1 and 2, which comes from the Farmer Jean example in lecture. This table should exist OUTSIDE any `r` code chunks in order to format properly since it's markdown (html in this case) and not R. Here, we use a pair of '\$'s to enter/exit math mode (this is LaTeX, for those of you who are familiar with and which I used to produce the lecture handouts), which allows us to type symbols like $X_1$, $\leq$ for the "less than or equals" sign and $\geq$ for the "greater than or equals" sign. Use $\,$ (thin space), $\;$ (medium space), $\quad$ (large space, equivalent to curent font size), $\qquad$ (two large spaces) for spacing, so you can leave a blank for decision variables with coefficients of 0. 
(Note: don't put two \$'s next to each other. Always put a space in between.). 

Maximize total profit using decision variables $X_1$, $X_2$ | Profit = 0.15 $X_1$ + 0.40 $X_2$
--- | --- 
Subject to |  
Budget Constraint | 0.20$X_1$ + 0.70$X_2$ $\leq$ 100
Space Constraint | $X_1$ + $X_2$ $\leq$ 200
Non-Negativity Constraint 1 | $X_1$ + $\quad$ $\geq$ 0
Non-Negativity Constraint 2 | $\quad$ + $X_2$ $\geq$ 0

You may also refer to: https://github.com/adam-p/markdown-here/wiki/Markdown-Cheatsheet#tables for additional information regarding table formatting. From the professor's experience, it takes a while to get used to making tables in Markdown, and even minor changes may break the table. The most common mistakes are: 

- not having a blank line before and a blank line after each table, 
- not having the correct number of \|'s, 
- not ending Math mode correctly, and 
- putting two \$s next to each other. 

The professor will not entertain emails regarding table formatting questions. Also note that for this assignment, please *IGNORE* integer requirements, i.e., just use real number (e.g. fractional answers) if/when they come up. We shall leave it to the last tutorial assignment for integer optimization.

## Part One: Lab Session Completion and Discussion

### Question 1 

The examples we discussed in lecture were all maximization problems (specifically, to maximize profit). In this question we shall explore minimization.

FunToys is famous for three types of toys: Cars, Animals, and Robots. Each year, near the holiday season, it receives large bulk orders for these items. To meet these orders, FunToys operates three small toy-making factories, A, B and C. 

- Factory A costs \$1000 per day to operate, and can produce 30 cars, 20 animals and 30 robots per day.
- Factory B costs \$1200 per day to operate, and can produce 40 cars, 50 animals and 10 robots per day.
- Factory C costs \$1500 per day to operate, and can produce 50 cars, 40 animals and 15 robots per day.

This Christmas, FunToys is required to deliver 5000 cars, 3000 animals and 2500 robots. You are tasked with finding out what is the most cost-efficient way to meet the order.


(1a) Let us start by trying to formulate this as a optimisation problem. 

- First, write down what you want to minimize.
- Second, write down your decision variables. What are you actually choosing?
- Third, write your objective function in terms of your decision variables.
- Fourth, write down the constraints: what are the contractual requirements you need to fulfill. What other constraints are there? Write them down in terms of your decision variables.
- Summarize them nicely in a table.


<p style="color:red">
First, I want to minimise total cost.
</p>

<p style="color:red">
Second, the decision variables are the numbers of days Factories A, B, and C operates respectively.
</p>

<p style="color:red">
Third, Cost = 1000 $X_1$ + 1200 $X_2$ + 1500 $X_3$.
</p>

<p style="color:red">
Fourth, the constraints are orders for cars, animals, and robots, and non-negativity constraints for the three variables.
</p>

<p style="color:red">
Table:
</p>

Minimise total cost using decision variables $X_1$, $X_2$, $X_3$ = number of days to run A, B, C respectively | Cost = 1000 $X_1$ + 1200 $X_2$ + 1500 $X_3$
--- | --- 
Subject to |  
Contract for cars | 30$X_1$ + 40$X_2$ + 50$X_3$ $\geq$ 5000
Contract for animals | 20$X_1$ + 50$X_2$ + 40$X_3$ $\geq$ 3000
Contract for robots | 30$X_1$ + 10$X_2$ + 15$X_3$ $\geq$ 2500
Non-Negativity Constraint 1 | $X_1$ + $\quad$ + $\quad$ $\geq$ 0
Non-Negativity Constraint 2 | $\quad$ + $X_2$ + $\quad$ $\geq$ 0
Non-Negativity Constraint 3 | $\quad$ + $\quad$ + $X_3$ $\geq$ 0

(1b) Write code to solve this optimisation problem. Report the optimal solution, and the value of the objective function at that solution. Interpret the solution: what do these numbers mean? (again, please ignore any integer requirements and just report fractional answers if they appear)


<p style="color:red">
Factory A 47.61905 days, Factory B 0 day, and Factory C 71.42857 days. The minimum total cost is 154761.9 dollars.
</p>

<p style="color:blue">
Optimal solution is: X1=47.6, X2=0.00, X3=71.43. That is, Factory A for 47.6 days, Factory B for 0 days and Factory C for 71.43 days. The Minimum cost is $154761.90. (1000x47.6)+(1200x0)+(1500x71.4)=154761.90.
</p>

```{r q1b-lp, echo=TRUE}
# type your code here
objective_function1 = c(1000, 1200, 1500)
constraint_mat1 = matrix(c(30, 40, 50, 20, 50, 40, 30, 10, 15),
                         ncol = 3,
                         byrow=TRUE)
constraint_dir1 = c(">=", ">=", ">=")
constraint_rhs1 = c(5000, 3000, 2500)

lp_solution1 = lp(direction = "min",
                  objective_function1,
                  constraint_mat1,
                  constraint_dir1,
                  constraint_rhs1,
                  compute.sens=TRUE)

print(lp_solution1$solution)

print(lp_solution1)
```

(1c) What if we impose an additional constraint that FunToys only has 60 days to complete the order? (Note that we can run all three factories *simultaneously*). What happens now?

Re-produce a new table summarizing the optimisation problem (including the existing and new constraints), and write R code to solve it. What is the new solution, and what is the objective function value?

<p style="color:red">Factory A 48.88889 days, Factory B 13.33333 day, and Factory C 60.00000 days. The minimum total cost is 154888.9 dollars.</p>

Minimise total cost using decision variables $X_1$, $X_2$, $X_3$ | Cost = 1000 $X_1$ + 1200 $X_2$ + 1500 $X_3$
--- | --- 
Subject to |  
Car Orders | 30$X_1$ + 40$X_2$ + 50$X_3$ $\geq$ 5000
Animal Orders | 20$X_1$ + 50$X_2$ + 40$X_3$ $\geq$ 3000
Robot Orders | 30$X_1$ + 10$X_2$ + 15$X_3$ $\geq$ 2500
Total Days 1 | $X_1$ + $\quad$ + $\quad$ $\leq$ 60
Total Days 2 | $\quad$ + $X_2$ + $\quad$ $\leq$ 60
Total Days 3 | $\quad$ + $\quad$ + $X_3$ $\leq$ 60
Non-Negativity Constraint 1 | $X_1$ + $\quad$ + $\quad$ $\geq$ 0
Non-Negativity Constraint 2 | $\quad$ + $X_2$ + $\quad$ $\geq$ 0
Non-Negativity Constraint 3 | $\quad$ + $\quad$ + $X_3$ $\geq$ 0

```{r q1c-lp, echo=TRUE}
# type your code here
objective_function2 = c(1000, 1200, 1500)
constraint_mat2 = matrix(c(30, 40, 50, 20, 50, 40, 30, 10, 15, 1, 0, 0, 0, 1, 0, 0, 0, 1),
                         ncol = 3,
                         byrow=TRUE)
constraint_dir2 = c(">=", ">=", ">=", "<=", "<=", "<=")
constraint_rhs2 = c(5000, 3000, 2500, 60, 60, 60)

lp_solution2 = lp(direction = "min",
                  objective_function2,
                  constraint_mat2,
                  constraint_dir2,
                  constraint_rhs2,
                  compute.sens=TRUE)

print(lp_solution2$solution)

print(lp_solution2)
```


(1d) For the solution in 1c, which of the constraints are binding, and which are non-binding?

<p style="color:red">
The binding constraints are Car Orders, Robot Orders, and Total Day Constraint 3.
</p>

<p style="color:blue">
Binding constraints are the Contract for cars, Contract for robots, and Time constraint for Factory C.
</p>

<p style="color:blue">
Nonbinding constraints are Contract for animals, Time constraints on Factories B and C, and All non-negativity constraints.
</p>

```{r q1d-lp, echo=TRUE}
# type your code here
print(lp_solution2$duals)

## my original answer
x1_2 <- lp_solution2$solution[1]
x2_2 <- lp_solution2$solution[2]
x3_2 <- lp_solution2$solution[3]

#car orders
print(30*x1_2 + 40*x2_2 + 50*x3_2 == 5000)

#animal orders
print(20*x1_2 + 50*x2_2 + 40*x3_2 == 3000)

#robot orders
print(30*x1_2 + 10*x2_2 + 15*x3_2 == 2500)

#total days 1
print(x1_2 == 60)

#total days 2
print(x2_2 == 60)

#total days 3
print(x3_2 == 60)

#non-negativity constraint 1
print(x1_2 == 0)

#non-negativity constraint 2
print(x2_2 == 0)

#non-negativity constraint 3
print(x3_2 == 0)
```


(1e) Using your solution in 1c, print out the Shadow Prices. Interpret these values -- make sure you can explain why each shadow price is zero or why it is positive/negative! Your answer from part d) should also help you explain. (Note again, that we IGNORE integer requirements in this question, so your $X$ variables can be fractional...)

<p style="color:blue">

</p>

```{r q1e-lp, echo=TRUE}
# type your code here
print(lp_solution2$duals)
```



## Part Two: Assignment Submission 

### Question 2 (total 20 points)

You work as a chief consultant for a family-owned Swiss watch workshop, producing three types of watches: casual watch, dress watch and luxury watch. Watch production mainly involves machine and labor inputs, in terms of machine and labor hours, respectively. For workshop's daily operation, you know that:

- Each casual watch brings in average profit of \$6k but production of each casual watch requires two hours on machine and half an hour of handcraft.
- Each dress watch brings in average profit of \$15k but production requires one and a half machine hours and two labor hours.
- Each luxury watch has an average profit of \$23k but production of luxury watch needs one hour of metalwork and four hours of handcraft. 

Workshop has two machines on metalwork (unbreakable German machine which is running 24 hours) and 6 watchmakers (each working on 12 hour shift daily). The workshop would like to know how to maximize their daily profit by deciding their production plan.

(2a) Clearly define the decision variables, and write out the objective function and constraints (using the table template provided above). Please include the non-negativity constraints. (4 points)

<p style="color:red">**Type here your answer here.**</p>


(2b) Program your optimization problem and solve the formulated linear problem above, using `lpSolve` in R. Is there any feasible solution and if yes, what is the optimal production plan to maximize workshop's daily profit? What is the daily profit at the optimal production plan? (4 points)

<p style="color:red">**Type here your answer here.**</p>
```{r q2b-lp-primal, echo=TRUE}
# type your code here
```



(2c) For some "what if" questions:

- What would be the largest profit margin of luxury watch such that it is optimal to not produce them in the workshop? (2 points)
- What would the least profit margin of casual watch such that it is optimal to produce them? (2 points)
- Given the fact that the constraining resources for further watch production expansion are machine and labor hour (as you may guess they are quite costly to procure: precision machining and experienced artisans), what would be the break-even procurement prices for one additional machine hour and one additional labor hour, respectively? (2 points)

<p style="color:red">**Type here your answer here.**</p>

```{r q2c-sens, echo=TRUE}
# type your code here
```

(2d) From the lecture you know that the shadow prices are the *internal prices* for the constrained resources, here in this case the machine and labor hours for the watch workshop. Let's now try to understand better how shadow prices are formed and why they are called `duals` in `lpSolve`. Consider a related but separate linear optimization problem.

Consider that workshop owner does not actually own the machine and labor capacities but need to rent them in the market by paying a pair of resource prices $(y_1, y_2)$ for machine hour and labor hour, respectively. She fixed a plan to rent two machines and six watchmakers to produce casual, dress and luxury watches same as before. She *decides* the resource prices $y_1$ and $y_2$ to pay for the resources she planned as to minimize daily renting cost, i.e. $48 y_1 + 72 y_2$. However, the constraints she faces now are that the marginal cost of producing each type of the watch must be as least as the profit margin of that watch (otherwise, she should not plan her production at the first place simply because she is losing money by producing watches). For example, for casual watch, $2y_1 + \frac{1}{2}y_2 \ge 6$ (production of each casual watch requires two hours on machine and half an hour of handcraft, marginal cost to produce one watch is thus $2y_1 + \frac{1}{2}y_2$).  

- Write out the decision variables, objective function and linear constraints of this minimization problem as what you did in (2a). Don't forget to include non-negativity constraints. (2 point)

- Code it into R and solve it. What is the optimal solution for this minimization problem? What is the value of objective function, i.e. daily renting cost at the optimal solution? What are the shadow prices? Compared with profit maximization problem before, what do you find? (3 point)


<p style="color:red">**Type here your answer here.**</p>

```{r q2d-dual, echo=TRUE}
# type your code here
```

(2e) Successfully knitting the RMD to HTML. (1 point)